Ignition of Fuel Beds by Firebrands

### 2022-05-09 introduction.tex
Updates to introduction and background

\chapter{Introduction}
\label{part:intro}

\blindtext

   Wildfires are a common and often necessary occurrence in many ecosystems frequently producing beneficial effects that lead to a healthier ecosystem. However, interactions of wildfires with the built environment are often disastrous, reducing homes to rubble and uprooting lives. In recent years wildfires have become more severe and resulted in an increase in homes lost. In order to understand the cause of the increase in home loss the mode by which homes are ignited needs addressed. One significant mode of structure ignition is by firebrands. A firebrand, alternatively called an ember, is a hot combusting particle of biomass that travels from an active fire to another surface or biomass (e.g., pine needles, leaves, or a crevice in a deck). If the firebrands have sufficient energy they  may ignite the material they land on, leading to ignition and potentially the destruction of a structure. Characterization of firebrands typically involves three parts: the generation of a firebrand in the fire itself, transport of a firebrand from the fire to the eventual landing point, and the interaction of the firebrand with the surrounding material. The overall objective of this work is to identify the parameters and processes that control the ignition of a fuel bed when an ember lands on it. Four specific objectives were identified to act as framework for achieving the overall goal based on the current knowledge of fuel bed ignition. The specific objectives of this work are as follows:
        \begin{enumerate}
            \item Determine the effects of particle morphology on ignition propensity
            \item Ascertain ignition dependence on heating location(s), mode, and rate
            \item Identify and quantify the influence of environmental conditions on ignition propensity.
            \item Identify the influence of fuel bed chemical composition on flaming ignition.
        \end{enumerate}
    The structure of this dissertation is as follows. First the current state of knowledge of fuel bed ignition is summarized as it applies to this work. The results of this effort are then presented in manuscript form, followed by the conclusions, suggestions for future work, and appendices. 
    The first manuscript addresses the effect of particle size on the ignition probability as well as the effects of heating rate resulting from changes in particle morphology. The second manuscript addresses the influence of the chemical composition on ignition propensity and how changing material properties effect the heat transfer both within the bed and between the bed and the ember. The third manuscript addresses the impact of environmental conditions, namely wind, on the ignition probability of fuel beds.

    Hopefully, the knowledge and conclusions of this work may be used to inform fire prevention and suppression efforts by creating more informed.

>>>>

\chapter{Introduction}
\label{part:intro}
    Wildfires are an inevitable and often beneficial occurrence in many ecosystems around the globe. However, wildfires are also a significant source of destruction especially when they transition from ecosystems that are adapted to them to the built environment where homes, structures, and sometimes whole communities are consumed.  In recent years the number of homes and structures lost to wildfires has increased and is likely to further increase due to three factors. Climate change is currently and anticipated to be a driving factor for increased fire severity~\cite{Levin2021Unveiling2019/2020}. Wildfire exclusion has led to a significant changes in ecosystems, promoting more severe fires~\cite{Marlon2012Long-termUSA, Keeley2019} and the expansion of the Wildland Urban Interface (WUI) has increased the number of homes in the path of fires~\cite{Radeloff2018RapidRisk, Hammer2009DemographicManagement, }. While the risk of loss due to fires is likely to increase for the foreseeable future, adapting structures to the increased risk by accurately predicting risks both before and during fires may reduce the number of structures lost~\cite{Manzello2021}.

    One significant mode of structure ignition is by firebrands~\cite{Manzello2020}. A firebrand, alternatively called an ember, is a hot combusting particle of biomass that travels from an active fire to another surface or biomass (e.g., pine needles, leaves, or a crevice in a deck). If the firebrands have sufficient energy they  may ignite the material they land on, leading to ignition, spot fires, and potentially the destruction of a structure. Characterization of firebrands typically involves three parts: the generation of a firebrand in the fire itself, transport of a firebrand from the fire to the eventual landing point, and the interaction of the firebrand with the surrounding material~\cite{Babrauskas2003}.

    The overall goal of this work is to identify parameters and processes that control the ignition of a fuel bed when an ember lands on it. Four specific objectives were addressed to act as framework for achieving the overall goal based on the current knowledge of fuel bed ignition. The specific objectives of this work are as follows:
        \begin{enumerate}
            \item Determine the effects of sizes of the recipient fuels on ignition behavior
            \item Ascertain ignition dependence on heating location(s), mode, and rate
            \item Identify and quantify the influence of wind on ignition propensity.
            \item Identify the influence of fuel bed chemical composition on flaming ignition.
        \end{enumerate}
    The structure of this dissertation is as follows. First the current state of knowledge of fuel bed ignition modelling is summarized as it applies to this work (Chapter~\ref{part:literature}). Literature review specific to each objective is contained in the corresponding manuscripts. The results of this effort are then presented in manuscript form, followed by the conclusions, suggestions for future work, and appendices. The first manuscript addresses sensitivities of ignition to the particle size of the fuel bed materials (Chapter~\ref{part:manuscript1}). The second manuscript explores the influence of wind and fluid phenomena around the embers in the presence of a heat source unaffected by wind (Chapter~\ref{part:manuscript2}). The third manuscript identifies sensitivities to ignition related to chemical composition and thermal properties in both quiescent conditions and with wind (Chapter~\ref{part:manuscript3}). The fourth manuscript identifies changes in ignition propensity when multiple embers are located in close proximity on the fuel bed (Chapter~\ref{part:manuscript1}). The final chapter summarizes the conclusions of each manuscript, discusses the implications of the conclusions with respect to predictive using an example ignition model, and suggests avenues for future work (Chapter~\ref{part:conclusion}).

    It is anticipated that this work will provide insights and a basis for the creation of a simplified model that can accurately predict ignition. For example, in a recent conversation with a developer of the Fire Dynamics Simulator (FDS) the need was expressed for an accurate and low computational cost methodology or model to predict ignition fro better predictions of fire spread. Due to the large scales in fire simulations (~10\si{\meter} grid size) a detailed model of ignition (<1\si{\milli\meter} grid size) is not feasible. Similarly, risk management personnel or homeowners are not likely to have the capability to create detailed models and predictions of risk around structures and homes they are tasked with protecting. The goal and perceived impact of this work is to create or enable the creation of models that can be effectively implemented by the people interacting with fires and their risks in the field.

### 2022-05-09 litreview.tex
Updates to introduction and background

\chapter{Background}
\label{part:literature}
    This chapter provides a discussion of the background and motivation of this work in further depth than the manuscripts in this dissertation.  Here, a general overview of ignitions as a threat to structures and fire spread as well as the current state of knowledge and predictive capabilities is communicated. Each manuscript includes a relevant and more detailed literature review pertaining to the specific objective of the work.

    Wild fires have been an integral part of many ecosystems for millennia and are often beneficial or required for maintaining a healthy ecosystem. As beneficial as they may be to the natural enviornment wildfires pose significant risk to the built environment. As communities around the world have expanded communities outward into environments that support wildfires (e.g., forests and grasslands) the risk of destruction of structures by wildfires has increased. In the United States 91\% of the Wildland Urban Interface \nomenclature{WUI}{Wildland Urban Interface} growth of 189,000\si{\kilo\meter\squared} (larger than the state of Washington) development between 1990 and 2010 is occupied by homes~\cite{Radeloff2017} and has continued to grow on recent years. The increased development in theses areas means that an excess of 40 million homes are located in areas that may be threatened by a wildfire.

    In addition to the increase of homes in the WUI the intensity and severity of fires has increased. California is a prime example of the increased fire severity where 13 of the 20 most destructive fires have occurred since 2010 and 7 of the 20 most destructive occurred in 2020 and 2021~\cite{CALFIRE2018}. Increasingly severe fires are not just limited to California or the United States. Texas and Tennessee each had one of the top ten most severe fires in the United States between 2005 and 2020 and Colorado, Oregon, Washington, Oklahoma, and Florida have also seen highly destructive fires in recent years~\cite{Barrett2020}. The United States is not unique in the presence of wilfdires that destroy homes. Australia's Black Summer in 2019-2020 was one of the most severe in the countries history~\cite{Levin2021}. The Mediterranean countries of Portugal, Spain, France, Italy, and Greece are also either experiencing greater losses or are a risk for greater losses due to wildfires in recent years. Similar trends are also present in many Asian and South American countries~\cite{Manzello2018}. Wildfires that threaten lives, homes, and structures are truly a global phenomena and it is essential to understand how homes are ignited and destroyed in wildfires to better prevent losses from occurring.

    Ignition of a home in a WUI fire is attributed to three primary mechanisms; direct flame contact, ignition by radiative heating, and ignition by firebrands. Of these methods ignition by firebrands is considered to be a primary threat to homes~\cite{Suzuki2021, Mell2010}. A case study of structure ignitions in a WUI fire found that firebrand ignition was responsible for 2/3 of structure ignitions~\cite{Mell2011}. With the majority of ignitions occurring due to firebrand ignition it is imperative to understand the mechanism of firebrand ignition to reduce home losses. The general mechanism for firebrand ignition is considered to be a three step process which starts within the fire itself, often at a significant distance from the ignition point. First, a firebrand is generated as burning material breaks off or is lofted into the air. The firebrand is then lofted through the air and transported to the ignition site. Upon landing on a recipient fuel bed ignition occurs if the criteria for ignition is met~\cite{Babrauskas2003}. When considering the possibility of ignition during a fire event all three of processes must be considered as each process significantly influences the properties of an ember landing on a fuel bed. Due to the highly variable nature of winds that transport firebrands during fires the firebrand the firebrand generation and fuel bed ignition processes may be the most consistent when considering prevention approaches. Foliage types are much more consistent and measureable than wind and have a significant impact on the type of embers produced~\cite{Hudson2020EffectsScale} and fuel beds are largely controlled and/or modifiable by homeowners. Unfortunately, the generation and ignition processes are significantly less understood than the transport processes~\cite{Manzello2020}. The work contained in this dissertation is focused on the ignition process of the fuel bed by a firebrand and does not consider the generation of firebrands. For further information on firebrand generation processes the reader is referred to review papers considering this process by Fernandez-Pello~\cite{Fernandez-Pello2017} and by Manzello et al.~\cite{Manzello2020}.

    Developing and increased understanding of the ignition of a fuel bed has been identified as a necessary step to better protect structures from firebrand ignition for decades but research was limited until the mid 2000's. Early research and a significant portion of current research has been conducted by dropping combusting firebrands (either flaming or glowing) or hot metal particles fuel beds in varying configurations that simulate conditions in a wildfire and observing ignition or no ignition outcomes. These works have identified some parameters that appear to control ignition but models created from these studies only provide qualitative predictions of ignition. The lack of models of quantitative accuracy was suggested by Finney et al. in 2013 who identified the lack of a fundamental understanding of what processes occur and how they interact as a primary gap in the knowledge of ignition~\cite{Finney2013}. A similar need was identified by Manzello et al. in 2018~\cite{Manzello2018} and as recently as 2020~\cite{Manzello2020}. As the fires increase in severity and the number of homes is likely to keep increasing in the future the need for a better understanding of the processes the control ignition has never been more urgent.

    In order to identify the controlling parameters of ignition and create a model of quantitative accuracy the processes that occur leading up to ignition must be identified. Figure~\ref{fig:ignitionDiagram} illustrates the different factors that can influence the ignition process. To better understand how these parameters effect the ignition process can be considered as a series of three processes.  The first is the heat transfer to the fuel bed from the ember. The second is the pyrolysis of the fuel bed material and the subsequent release of pyrolyzates into the air above the fuel bed. The third step is the mixing of the pyrolyzates with air above the fuel bed and subsequent gas phase ignition. It is important to note that all three of these processes may occur simultaneously throughout the ignition process and are often interconnected. For example, as heat is transferred from the firebrand to the fuel bed the thermal conductivity of the fuel bed changes with temperature both before and after pyrolysis begins~\cite{Fjellerup2003}. Furthermore, the contact area between the firebrand and fuel continuously changes as mass is released due to pyrolysis. Due to the interconnection of the processes that occur during ignition and the inherent variability of fuel beds and biomass fuels in general quantification of parameters that influence ignition is quite challenging. Despite these challenges results of previous research have identified influential parameters that have consistently been observed to control ignition. This work considers these parameters as belonging to one of three categories namely, thermo-physical properties of the fuel bed, energy supplied by the firebrand, and environmental conditions near the ignition point. Detailed reviews of each category are provided in the relevant manuscripts and are not reproduced here. The remaining sections of this review address the predictive models of ignition that have been created from the current state of knowledge. The accuracy and appilicability of each model is discussed in turn and the remaining challenges are identified as they have motivated the objectives of this work. 
        \begin{figure}[hpbt]
            \centering
                \includegraphics[width=0.75\textwidth]{Figures/dissertationBed.pdf}
            \caption{Diagram illustrating potential parameters that may influence ignition of a fuel bed.}
            \label{fig:ignitionDiagram}
        \end{figure}

@@

        \begin{equation}
            r_{cr} = \delta_{cr}\sqrt{\frac{k}{\rho A \Delta H}\frac{RT^{2}_{p0}}{E} exp\left(\frac{E}{RT_{p0}} \right)}
            \label{eqn:hotspot}
        \end{equation}
    In this equation $\delta_{cr}$ is defined as the critical Frank-Kamenetskii hot spot parameter as determined from calculations~\cite{Goldshlager} and is dependent on the ratio of volumetric heat capacities of the ignition source and the material, the activation temperature of the material, and the temperature of the hot particle. Ultimately, Equation~\ref{eqn:hotspot} defines the particle energy needed to raise the temperature of a fuel bed region above the activation temperature to enable thermal runaway and subsequent ignition. While a relatively simple, the hot spot theory has some limitations that prevent quantitative predictions of ignition in a wildfire setting. A discussion of the shortcomings of the hot spot theory are presented by Thomas~\cite{Thomas1965} and more recently presented by Hadden et al.~\cite{Hadden2011} with respect to wildfires. The two most significant shortcomings addressed by the previous works are neglecting heat losses to the surroundings and considering the pyrolysis and ignition phenomena as single step infinitely fast reactions. Hadden et al. attributes these shortcomings as a significant source of error for quantitative ignition predictions. This assertion is reinforced in a review by Fernandez-Pello who concluded that in instances where hot spot theory has been applied only qualitatively accurate predictions of ignition have been obtained. Ultimately, the hot spot theory does not capture the complexity and variability inherent to ignition.

@@

    Proposed ignition models that provide greater accuracy than the hot spot and critical mass flux models do so at the expense of complexity and computational cost. The majority of these models and modelling efforts use a two part approach to join a pyrolysis specific solver to address the chemical kinetics of pyrolysis to an existing or custom computation fluid dynamics solver. This approach requires selecting choosing a combination of three parts each of which has limitations and considerations which may affect the accuracy of ignition predictions. First an appropriate chemical kinetic mechanism must be chosen based on the computational resources available. For example, many early works~\cite{DiBlasi1996} and some more recently~\cite{Ding2015, Torero2020} have considered pyrolysis occurring as a single step process or as a process of few reactions (approximately ten or fewer). Newer kinetic mechanisms have been developed to more completely capture the processes of pyrolysis and are much more complex. Perhaps the most complete mechanisms are extended forms of the mechanism developed by Ranzi et al.~\cite{Ranzi2008}. Additions to this mechanism have expanded the method to include extractives to cover a greater variety of biomass~\cite{Debiagi2015} and the addition of reactions to address gas phase reactions previously not considered~\cite{Dhahak2019}.

    One of the most recent, and most promising, tools for ignition modeling and pyrolysis is Gpyro which is a pyrolysis solver specifically designed ~\cite{Lautenberger2009a}. Unlike other

    Unfortunately, this model and models like it are too time consuming given current computational resources to be applied in across the vast number of conditions and configurations of possible ignition threats in wildland urban interface regions around the globe. Thus, a useable and accurate model will likely land somewhere between the critical mass flux and full CFD model with respect to complexity. Currently, the understanding of which effects are the most influential to ignition is not complete enough to develop such a model. Bridging this knowledge gap by identifying an quantifying which parameters are the most influential to an ignition or non-ignition outcome is the overall goal of this work.

    A significant number of ignition criteria and models have been proposed to predict fuel bed ignition, however no model has provided a general quantitatively accurate prediction of ignition. 
    \section{Current Ignition Models}
        \begin{itemize}
            \item Hot spot theory (Hadden 2011) (Done)
            \item Critical mass flux (Nelson 1995)
            \item Energy content of the particle
            \item Numerical methods
            \item Analytic correlations
        \end{itemize}

>>>>

\chapter{Background}
\label{part:literature}
    This chapter provides a discussion of the background and motivation of this work in further depth than the manuscripts presented later in this dissertation.  Here, a general overview of ignitions as a threat to structures and fire spread as well as the current state of knowledge and predictive capabilities is communicated. Each manuscript includes a relevant and more detailed literature review pertaining to the specific objective of the chapter.

\section{Widlfires and the WUI}
    Wildfires have been an integral part of many ecosystems for millennia and are often beneficial or required for maintaining a healthy ecosystem. As beneficial as they may be to the natural enviornment wildfires pose significant risk to the built environment. As communities around the world have expanded communities outward into environments that support wildfires (e.g., forests and grasslands) the risk of destruction of structures by wildfires has increased~\cite{Hammer2009}. In the United States, 91\% of the Wildland Urban Interface (WUI) \nomenclature{WUI}{Wildland Urban Interface} growth of 189,000\si{\kilo\meter\squared} (larger than the state of Washington) development between 1990 and 2010 is occupied by homes~\cite{Radeloff2017} and has continued to grow on recent years. The increased development in theses areas means that an excess of 40 million homes are located in areas that may be threatened by a wildfire.

    In addition to the increase of homes in the WUI the intensity and severity of fires has increased. California is an example of the increased fire severity where 13 of the 20 most destructive fires have occurred since 2010 and 7 of the 20 most destructive occurred in 2020 and 2021~\cite{CALFIRE2018}. Increasingly severe fires are not just limited to California. Texas and Tennessee each had one of the top ten most severe fires in the United States between 2005 and 2020. Colorado, Oregon, Washington, Oklahoma, and Florida have also seen highly destructive fires in recent years~\cite{Barrett2020}. The United States is not unique in the presence of wilfdires that destroy homes. Australia's Black Summer in 2019-2020 was one of the most severe fire seasons in the countries' history~\cite{Levin2021Unveiling2019/2020} where over 3000 homes and greater than 46 million acres  burned~\cite{Filkov2020}, which is approximately equivalent to 75\% of Oregon's land area. The Mediterranean countries of Portugal, Spain, France, Italy, and Greece are also either experiencing greater losses or are a risk for greater losses due to wildfires in recent years. Similar trends are also present in many Asian and South American countries~\cite{Manzello2018}. Wildfires that threaten lives, homes, and structures are truly a global phenomena and it is essential to understand how homes are ignited and destroyed in wildfires to better prevent losses from occurring.

\section{The Problem of Ignition}
    Ignition of a home in a WUI fire is attributed to three primary mechanisms; direct flame contact, ignition by radiative heating, and ignition by firebrands. Of these methods ignition by firebrands is considered to be a primary threat to homes~\cite{Suzuki2021, Mell2010, Manzello2018FORUMResearch}. For example, an evaluation of structural ignitions in a WUI fire found that firebrand ignition was responsible for 2/3 of structure ignitions~\cite{Mell2011}. With the majority of ignitions occurring due to firebrand ignition it is imperative to understand the mechanism of firebrand ignition to reduce home losses. Codes, standards, and best practices built upon a better understanding of ignition by firebrands would help provide better resilience to firebrand attacks~\cite{Manzello2020} reducing the number of homes lost in wildfires. Examples of these changes may include adjustments in building and landscaping materials favoring the selection of harder to ignite species or housing geometry constraints to reduce collection of flammable materials or embers. Without a comprehensive knowledge of ignition setting codes and standards that reduce risk of home loss is challenging at best and may even be counterproductive.

    The general mechanism for firebrand ignition is considered to be a three step process which starts within the fire itself, often at a significant distance from the ignition point. First, a firebrand is generated as burning material breaks off or is lofted into the air. Second, the firebrand is then lofted through the air and transported to the ignition site. Third, a firebrand lands on a recipient fuel bed and then ignition occurs if the criteria for ignition is met~\cite{Babrauskas2003}. When considering the possibility of ignition during a fire event all three of processes must be considered as each process significantly influences the properties of an ember landing on a fuel bed. For example, the number, energy content, and size of embers that may land on a fuel bed is significantly influence by the surrounding foliage~\cite{Hudson2020EffectsScale} and moisture content of the foliage~\cite{Adusumilli2021FirebrandShrub} which introduces significant variability in ignition sources a fuel bed may encounter. Similarly, the winds transporting a firebrand significantly influence the path of travel and time of flight which influences the energy content~\cite{Sardoy2007, MatvienkoOVA2016} and localized winds around structures may lead to firebrand accumulation~\cite{Suzuki2020a, Suzuki2017, Suzuki2015}. Unfortunately, the generation and ignition processes are significantly less understood than the transport processes~\cite{Manzello2020}. The work contained in this dissertation is focused on the ignition process of the fuel bed by a firebrand and does not consider the generation of firebrands. For further information on firebrand generation processes the reader is referred to review papers considering this process by Fernandez-Pello~\cite{Fernandez-Pello2017} and by Manzello et al.~\cite{Manzello2020}.

    Ignition of a fuel by firebrands has been identified as a source of house loss f
    Developing an increased understanding of the ignition of a fuel bed has been identified as a necessary step to better protect structures from firebrand ignition since at  ~\cite{Waterman1969ExperiemntalGeneration}. but research was limited until the mid 2000's.

    Early research and a significant portion of current research has been conducted by dropping combusting firebrands (either flaming or glowing) or hot metal particles fuel beds in varying configurations that simulate conditions in a wildfire and observing ignition or no ignition outcomes. These works have identified some parameters that appear to control ignition but models created from these studies only provide qualitative predictions of ignition. The lack of models of quantitative accuracy was suggested by Finney et al. in 2013 who identified the lack of a fundamental understanding of what processes occur and how they interact as a primary gap in the knowledge of ignition~\cite{Finney2013}. A similar need was identified by Manzello et al. in 2018~\cite{Manzello2018} and again in 2020~\cite{Manzello2020}. As fires increase in severity, and the WUI continues to expand the number of homes lost to wildfires is likely to significantly increase in the near future. Thus, the need for accurate predictions of ignition has never been higher.

\section{Modeling Efforts}
    Creating a model with quantitative accuracy of ignition requires a through understanding of the processes and relative importance of each processes leading up to ignition.    Figure~\ref{fig:ignitionDiagram} illustrates the different factors that can influence the ignition process. To better understand how these parameters effect the ignition process can be considered as a series of three processes.  The first is the heat transfer to the fuel bed from the ember. The second is the pyrolysis of the fuel bed material and the subsequent release of pyrolyzates into the air above the fuel bed. The third step is the mixing of the pyrolyzates with air above the fuel bed and subsequent gas phase ignition. It is important to note that all three of these processes may occur simultaneously throughout the ignition process and are often interconnected. For example, as heat is transferred from the firebrand to the fuel bed the thermal conductivity of the fuel bed changes with temperature both before and after pyrolysis begins~\cite{Fjellerup2003}. Furthermore, the contact area between the firebrand and fuel continuously changes as mass is released due to pyrolysis. Due to the interconnection of the processes that occur during ignition and the inherent variability of fuel beds and biomass fuels in general quantification of parameters that influence ignition is quite challenging. Despite these challenges results of previous research have identified influential parameters that have consistently been observed to control ignition. This work considers these parameters as belonging to one of three categories namely, thermo-physical properties of the fuel bed, energy supplied by the firebrand, and environmental conditions near the ignition point. Detailed reviews of each category are provided in the relevant manuscripts and are not reproduced here. The remaining sections of this review address the predictive models of ignition that have been created from the current state of knowledge. The accuracy and appilicability of each model is discussed in turn and the remaining challenges are identified as they have motivated the objectives of this work. 
        \begin{figure}[hpbt]
            \centering
                \includegraphics[width=0.75\textwidth]{Figures/dissertationBed.pdf}
            \caption{Diagram illustrating potential parameters that may influence ignition of a fuel bed.}
            \label{fig:ignitionDiagram}
        \end{figure}

@@

        \begin{equation}
            r_{cr} = \delta_{cr}\sqrt{\frac{k}{\rho A \Delta H}\frac{RT^{2}_{p0}}{E} exp\left(\frac{E}{RT_{p0}} \right)}
            \label{eqn:hotspot}
        \end{equation}
    In this equation $\delta_{cr}$ is defined as the critical Frank-Kamenetskii hot spot parameter as determined from calculations~\cite{Goldshleger1973IgnitionDimensions} and is dependent on the ratio of volumetric heat capacities of the ignition source and the material, the activation energy of the material, and the temperature of the hot particle. Ultimately, Equation~\ref{eqn:hotspot} defines the particle energy needed to raise the temperature of a fuel bed region such that enable thermal runaway and subsequent ignition occur. While a relatively simple, the hot spot theory has some limitations that prevent quantitative predictions of ignition in a wildfire setting. A discussion of the shortcomings of the hot spot theory are presented by Thomas~\cite{Thomas1965} and more recently presented by Hadden et al.~\cite{Hadden2011} with respect to wildfires. The two most significant shortcomings addressed by the previous works are neglecting heat losses to the surroundings and considering the pyrolysis and ignition phenomena as single step infinitely fast reactions. Hadden et al. attributes these shortcomings as a significant source of error for quantitative ignition predictions. This assertion is reinforced in a review by Fernandez-Pello~\cite{Fernandez-Pello2017} who concluded that in instances where hot spot theory has been applied only qualitatively accurate predictions of ignition have been obtained. A similar model to hot spot theory but with a more comprehensive inclusion of heat loss from the ember was developed by Yin et al.~\cite{Yin2014}. A correlation between fuel properties, energy transferred to the fuel bed, and the time to ignition was proposed and is shown in Equation~\ref{eqn:yinMC}.
        \begin{equation}
            \sqrt{t_{ig}} \approx \frac{q\sqrt{\frac{\rho_{f}k}{c_{p}}}}{\frac{\rho Z \Delta H_c}{t_b}- h_T(T_f - T_0)}
            \label{eqn:yinMC}
        \end{equation}
    Comparing the correlations presented in Equations~\ref{eqn:hotspot} and \ref{eqn:yinMC}, Equation~\ref{eqn:yinMC} includes properties of the fuel bed ($\rho$, $k$, $c_p$, $Z$, $T_0$) and properties the ember or energy source ($\rho_f$, $\Delta H_c/t_b$, $T_f$) as well as the energy required for ignition ($q$) mirroring the Frank-Kamenetskii critical hot spot diameter. The addition of heat losses to the ambient is the most significant difference from the hotspot parameter and addresses a key shortcoming of the hot spot theory. However, the model does not consider the combustion of the fuel bed other than the critical energy for ignition leaving the second key gap for the hot spot theory unaddressed. Perhaps due to the omission of combustion, this model suffers a similar fate as the hot spot theory and follows qualitative trends but is not quantitative. Ultimately, the hot spot theory and similar models do not capture the complexity and variability inherent to ignition~\cite{Manzello2020}.

@@

    Proposed ignition models that provide greater accuracy than the hot spot and critical mass flux models do so at the expense of complexity and computational cost. The majority of these models and modelling efforts using a two part approach to join a pyrolysis specific solver to address the chemical kinetics to an existing or custom computation fluid dynamics solver. This approach requires selecting choosing a combination of three parts each of which has limitations and considerations which may affect the accuracy of ignition predictions. First, an appropriate chemical kinetic mechanism must be chosen based on the computational resources available. For example, many early works~\cite{DiBlasi1996} and some more recently~\cite{Ding2015, Torero2020} have considered pyrolysis occurring as a single step process or as a process of few reactions (approximately ten or fewer). Newer kinetic mechanisms have been developed to more completely capture the processes of pyrolysis and are much more complex. Perhaps the most complete mechanisms are extended forms of the mechanism developed by Ranzi et al.~\cite{Ranzi2008}. Additions to this mechanism have expanded the method to include extractives to cover a greater variety of biomass~\cite{Debiagi2015} and the addition of reactions to address gas phase reactions previously not considered~\cite{Dhahak2019}. For a more detailed discussion of the progression and challenges of biomass pyrolysis the reader is referred to review articles by Di Blasi~\cite{DIBLASI199371} and more recently by Haberle et al.~\cite{Haberle2017}.

    One of the most recent, and most promising, tools for ignition modeling and pyrolysis is Gpyro which is a pyrolysis solver specifically designed handle both the condensed (solid) and gas phase of biomass combustion~\cite{Lautenberger2009a}. Gpyro is often coupled coupled with the Fire Dynamics Simulator (FDS)~\cite{FDSManual}, a LES solver specifically designed for physics relevant in fires, to form a complete computational package. The combination of Gpyro and FDS have been utilized to observed combustion sensitivities and ignition senstivities. For example a study of the effects of moisture content on ignition using Gpyro/FDS suggested that moisture in the fuel affected ignition, gas phase combustion, and pyrolysis of the fuel~\cite{Yashwanth2016}. Other studies using Gpyro/FDS have been conducted to predict ignition and observe controlling parameters~\cite{Shotorban2018, Fernandez-Pello2015, Lautenberger2009}. Unfortunately, the Gpyro/FDS modelling require significant computational expenditure and are not practical to be used for the prediction of ignition across the  ast number of conditions and configurations of possible ignition threats in wildland urban interface regions around the globe. Thus, a usable and accurate model will likely land somewhere between the critical mass flux and full CFD model with respect to complexity. Currently, the understanding of which effects are the most influential to ignition is not complete enough to develop such a model. Bridging this knowledge gap by identifying an quantifying which parameters are the most influential to an ignition or non-ignition outcome is the overall goal of this work.

\section{Summary}
